Probabilistic Forecasting of the Disturbance Storm Time Index: An Autoregressive Gaussian Process approach

### 2017-03-02 agutmpl.tex
Modifications Iteration: 1

After selecting the best performing GP-AR and GP-ARX models in the validation phase, we test and compare the performance of these models with the predictions generated from the \emph{persistence model} $\hat{D}st(t) = Dst(t-1)$, on a set of 63 storm events occurring between 1998 and 2006 as given in table \ref{table:teststorms}.

@@

Figures \ref{fig:CompareMae} and \ref{fig:CompareCC} show how the mean absolute error and coefficient of correlation as calculated on the validation set of table \ref{table:validationstorms}, vary with increasing model order for GP-AR and GP-ARX. In both cases the results first improve and then stagnate or worsen.

@@

Figures \ref{fig:ComparePred1}, \ref{fig:ComparePred2} and \ref{fig:ComparePred3} show OSA predictions of the GP-ARX model with $\pm \sigma$ error bars for three storm events in the years 1998 and 2000. The GP-ARX model gives accurate predictions along with plausible error bars around its mean predictions.

@@

Our results can be summarized as follows.
\begin{enumerate}
      \item \emph{Persistence} model in the model building and evaluation process in the context of \emph{one step ahead} prediction of the $Dst$ index. It is clear that the \emph{persistence} behavior in the $Dst$ values is very strong i.e. the trivial predictive model $\hat{Dst}(t) = Dst(t-1)$ gives excellent performance according to the metrics chosen.

@@

      \item Since the GP models generate predictive distributions for test data and not just point predictions they lend themselves to the requirements of space weather prediction very well because of the need to generate error bars on predictions.
\end{enumerate}

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Compare-mae.png}
\caption{Mean Absolute Error on validation set storms vs model order for GP-AR and GP-ARX}
\label{fig:CompareMae}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Compare-cc.png}
\caption{Mean Absolute Error on validation set storms vs model order for GP-AR and GP-ARX}
\label{fig:CompareCC}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Compare-mae-arx.png}
\caption{Mean Absolute Error on validation set storms vs model order for GP-AR and GP-ARX}
\label{fig:CompareMaeARX}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Compare-cc-arx.png}
\caption{Mean Absolute Error on validation set storms vs model order for GP-AR and GP-ARX}
\label{fig:CompareCCARX}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{PredictionsModel1/PredErrBars_Storm5.png}
\caption{OSA Predictions with $\pm \sigma$ error bars for event: 1998/09/25 01:00 to 1998/09/26 00:00}
\label{fig:ComparePred1}
\end{figure}

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth]{PredictionsModel1/PredErrBars_Storm14.png}
\caption{OSA Predictions with $\pm \sigma$ error bars for event: 2000/04/06 17:00 to 2000/04/08 09:00}
\label{fig:ComparePred3}
\end{figure}

>>>>

After selecting the best performing GP-AR and GP-ARX models in the validation phase, we test and compare the performance of these models with the predictions generated from the \emph{persistence model} $\hat{D}st(t) = Dst(t-1)$, on a set of 63 storm events occurring between 1998 and 2006 as given in table \ref{table:teststorms}, which is the same list of storm events as used in \citet{Ji2012}.

@@

Figures \ref{fig:CompareMae} and \ref{fig:CompareCC} show how the mean absolute error and coefficient of correlation as calculated on the validation set storm events of table \ref{table:validationstorms}, vary with increasing model order for GP-AR and GP-ARX. The results are represented as boxplots, in which a rectangle is drawn to represent the second and third quartiles, with a vertical line inside to indicate the median value while outlying points are shown as dots. In both cases, the predictive performance first improves and then stagnates or worsens with increasing model order.

@@

Figures \ref{fig:ComparePred1}, \ref{fig:ComparePred2} and \ref{fig:ComparePred3} show OSA predictions of the GP-ARX model with $\pm \sigma$ error bars for three storm events in the time period between 1998 and 2003. The GP-ARX model gives accurate predictions along with plausible error bars around its mean predictions.

@@

Our results can be summarized as follows.
\begin{enumerate}
      \item \emph{Persistence} model plays an important role in the model building and evaluation process in the context of \emph{one step ahead} prediction of the $Dst$ index. It is clear that the persistence behavior in the $Dst$ values is very strong i.e. the trivial predictive model $\hat{Dst}(t) = Dst(t-1)$ gives excellent performance according to the metrics chosen.

@@

      \item Since the GP models generate predictive distributions for test data and not just point predictions they lend themselves to the requirements of space weather prediction very well because of the need to generate error bars on predictions.

      \item The \emph{Gaussian Process} regression framework described in this study can also be extended to multiple hour ahead prediction of $Dst$, which is currently a work in progress.
\end{enumerate}

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Compare-mae.png}
\caption{Mean Absolute Error on validation set storms vs model order for GP-AR and GP-ARX. \\ \textbf{Key}: Rectangle borders represent the second and third quartiles, with a horizontal line inside to indicate the median value while outlying points are shown as dots}
\label{fig:CompareMae}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Compare-cc.png}
\caption{Coefficient of Correlation on validation set storms vs model order for GP-AR and GP-ARX \\ \textbf{Key}: Rectangle borders represent the second and third quartiles, with a horizontal line inside to indicate the median value while outlying points are shown as dots}
\label{fig:CompareCC}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Compare-mae-arx.png}
\caption{Mean Absolute Error on validation set storms vs model order for GP-AR and GP-ARX for \emph{CSA}, \emph{GS} and \emph{ML} model selection routines \\ Rectangle borders represent the second and third quartiles, with a horizontal line inside to indicate the median value while outlying points are shown as dots}
\label{fig:CompareMaeARX}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Compare-cc-arx.png}
\caption{Coefficient of Correlation on validation set storms vs model order for GP-AR and GP-ARX for \emph{CSA}, \emph{GS} and \emph{ML} model selection routines \\ Rectangle borders represent the second and third quartiles, with a horizontal line inside to indicate the median value while outlying points are shown as dots}
\label{fig:CompareCCARX}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{PredictionsModel1/PredErrBars_Storm43.png}
\caption{OSA Predictions with $\pm \sigma$ error bars for event: 2003/06/17 19:00 to 2003/06/19 03:00}
\label{fig:ComparePred1}
\end{figure}

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth]{PredictionsModel1/PredErrBars_Storm9.png}
\caption{OSA Predictions with $\pm \sigma$ error bars for event: 1999/01/13 16:00 to 1999/01/14 20:00}
\label{fig:ComparePred3}
\end{figure}
